### 2022-09-04 submission/reviewer_response.rmd
drafting changes based on reviewer comments

---
title: "Response to reviewers"
output:
  bookdown::pdf_document2: default
geometry: margin=1.0in
toc: true
font-size: 12pt
editor_options: 
  chunk_output_type: console
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \usepackage[left, pagewise]{lineno}
  - \usepackage{caption}
  - \usepackage{amsmath}
  - \doublespacing
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \date{}
  - \newcommand\given[1][]{\:#1\vert\:}
  - \newcommand{\indep}{\perp \!\!\! \perp}
bibliography: references.bib
csl: jce.csl
---

# Reviewer 1 {-}
"Individualized treatment effect was predicted best by modeling baseline risk in
interaction with treatment assignment"
Comments to the Authors:

I read the manuscript with interest, however, the authors first need to address
the following concerns.

- Page 2.  Line 6. Need to clarify what it means by baseline risk, as well as
what it means by the “reference.”  Line 10. Related to the comment above, can
you provide a sentence or two briefly describing what risk-modeling approaches
are?  Line 12. Need to clarify what it means by “absolute benefit.”
Line 13. Not clear, from the written sentence, which methods (whether they are
risk-based methods or the method based on modeling treatment effect-modifiers)
lead to the weaker discrimination of benefit in the presence of true effect
modifiers.  Line 18. The estimates refer to estimates of which parameters? Need
to clarify. Also, please indicate why these estimates may not be applied to
individual patients. \textcolor{green}{[DONE]}

- Page 4.  Line 1. Not clear what it means by the "AUC value of 0.75." Need to
explain AUC. And after all, what is the "true" value of $\beta$ that was used in the
simulations?  Line 6. What is c? And which term in the equation is the term that
corresponds to the constant treatment effect? What are the values of vs used in
the simulations?  Line 11. What do you mean by “standard” alpha? what is
"standard"? What is "effect"? Does it refer to the effect on the approaches'
performance? What do you mean by the effect of risk model discriminative
ability? How can risk model discriminative ability have any effects? Please
clarify. \textcolor{green}{[DONE]}

- Page 5.  Line 14. Please clarify how $\delta_0$, $\delta_1$ and $\delta_3$ are
  estimated. \textcolor{green}{[DONE]}

- Page 6.  Line 8. Please describe in a sentence what "ICI for benefit" is.
Line17. Please comment on how those 6 predictors were selected (is it also
selected based on AIC?). \textcolor{green}{[DONE]}

-  Page 7.  Line 14. Please describe how the "absolute benefit predictions" were
evaluated (please present the evaluation metric). \textcolor{green}{[DONE]}

- Page 8. In the figures, I think for each panel you should label the X-axis
(i.e., the methods under consideration). At first, it was difficult to
understand what was displayed. \textcolor{green}{[DONE]}

- Page 13. In Figure 4, again, what is the C-statistic for benefit? Need an
explicit expression or a clear description (both in figure caption and in the
text). \textcolor{green}{[DONE]}

- Page 16. In Figure 6, please also display the model-based confidence bands for
all the methods. Also please clarify how the predicted benefit was
computed. Does it just mean the treatment effect? If so, how could the "constant
treatment effect" method have a non-constant treatment effect as a function of
the baseline risk? Please clarify. \textcolor{green}{[DONE]}

- Page 18.  Line 21. Generally, in the estimation, how do you distinguish the
treatment-related harm from the actual treatment effect? \textcolor{red}{[UNRESOLVED]}

- Page 19.  Line 1. As discussed in the Discussion, the authors only considered
the risk-based methods. Along this line, I think the title of the manuscript is,
to some extent, misleading, because it gives an impression that the modeling
baseline risk-and-treatment interaction is a “globally best method” for the
individualized treatment effect estimation, which is not likely to be true.
Line 2. Why only tree-based approaches were cited? There is a large volume of
treatment effect modeling methods, which I think should be cited as well, if
tree-based approaches are to be cited.  Line 8. I don’t think what is stated in
line 8 is true when the "true" data model is defined based on separate
treatment-by-covariate interaction terms, in addition to the main effect
terms. Furthermore, there are many methods that avoid overfitting through
regularization and simultaneous treatment effect modifier selection. Authors
should discuss this. \textcolor{red}{[UNRESOLVED]}

# Reviewer 2 {-}

I reviewed the same manuscript last month. It seems that the manuscript is not
revised and is the same as the previous version and did not address any of the
comments. Please see my comments made in the previous round of review. Would you
please check if this attached manuscript is a revised version ? Thank you.

# Reviewer 3 {-}

The content of this paper is methodologically sound and the results
are consistent with what the literature has shown (namely that more complex
models are often outperformed by simpler ones in small sample and weaker signal
scenarios) and the writing is generally adequate. However, as written, the paper
does not focus on the significance of this methodology and the simulation
studies in terms of guiding clinical care (see Comment 1 for more details); this
was the main weakness of this paper.

1. The introduction will benefit from more emphasis on the practical
   significance of modeling baseline risk and treatment effects. In particular,
   in which real-life scenarios was this type of analysis useful and what
   happened when the baseline risk and its interaction with treatment was
   neglected in favor of a simpler approach? What type of analyses are typically
   done in such scenarios and what does the approach presented in this
   manuscript offer on top of the usual analysis methods?

2. During the application of restricted cubic splines, was there an assumption
   of interaction between risk and treatment effects as well? For example, were
   the restricted cubic splines fit separately to each treatment arm?

3. It would also be helpful to discuss the results (particularly the fact that
   the simpler model--linear interaction--outperforms RCS-3 in smaller sample
   size scenarios and weaker signals, even when there is nonlinearity in the
   true underlying relationship, as well as the fact that RCS-3 outperforms
   RCS-4 and RCS-5 in many scenarios) in terms of the bias-variance tradeoff. In
   particular, the restricted cubic splines-based methods introduce additional
   degrees of freedom and thus increase the variance, despite being expected to
   reduce the bias.

# Reviewer 4 {-}

This paper presents the results from a very large simulation study to
evaluate a number of fairly simple statistical approaches to estimate individual
treatment effects using data from a RCT. The paper is well written and the
results are presented objectively. The results show essentially what you might
expect, that if the data generating method is more complicated then flexible
methods perform best, while if the data generation method is more simple then
simpler methods do best. Furthermore, with larger sample size flexible methods
tend to do better. The authors end up essentially recommending an approach that
has a linear interaction term, or one with a cubic spline. A couple of the
comparators are even more simple than these (constant treatment effect and
stratification by quartiles), and it is no surprise that these naïve method are
not good in the simulations. So if the point of the paper is to discourage use
of these naïve methods, then that is a valuable contribution.  The big
limitation of the whole study is that the baseline risk (i.e. the linear
combination of the X variables from the control arm of the RCT) is assumed to be
the driver of the individual treatment effect. The data is generated with this
assumption, and all the estimation methods make this assumption. The authors
totally acknowledge this limitation. But I would have thought that in the
majority of realistic situations the linear combination of X’s for the main
effect would have been different from the linear combination of X’s in the
interaction with treatment. In this case the authors conclusion that linear
interaction and RCS3 are good, would likely not hold up. The authors do
reference a few articles that use tree-based methods (there are many many more
and lots of non tree-based approaches too), and correctly point out their
potential for overfitting. But regularization and cross-validation approaches
can do a pretty good job at limiting the overfitting. So I think I would favor
tree-based (or other more modern approaches), unless there is some clinical or
scientific rationale to believe that the control group risk is the driver of the
treatment effect heterogeneity.  A couple of others things that should be
mentioned as caveats. In the simulated data the X variables are made independent
of each other. If they were made to be correlated, that would probably be more
realistic. Do you have any information on how this might change the findings, in
general terms.  Larger sample sizes also opens up the possibility of
generalizing the linear additive assumption for the control group risk. For
example, this part of the model could now include non-linear terms in X and
interactions between X’s.